---
title: "R Notebook"
output: html_notebook
---


```{r, warning=FALSE, message=FALSE}
install.packages(c('ggthemes', 'scales'))
```

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggthemes)
library(scales)
donor <- read.csv('https://goo.gl/tm9JQ5')
police <- read.csv('https://goo.gl/nNAuDy')

donor %>% head(2)
# donor %>% ls()
# donor %>% summary()

# police %>% head(2)
# police %>% ls()
# police %>% summary()
```

```{r, warning=FALSE, message=FALSE}
mpg %>% ggplot() + geom_point(aes(displ, hwy))
```

```{r, warning=FALSE, message=FALSE}}
donor %>% ggplot(aes(primary_general)) + geom_bar()
```

now with color
```{r, warning=FALSE, message=FALSE}}
donor %>% ggplot(aes(primary_general, fill = primary_general)) + geom_bar()
```

exercise with police data
```{r, warning=FALSE, message=FALSE}}
police %>% 
  filter(
    district_sector %in% c('B', 'E', 'D', 'R', 'O', 'C', 'K') &
          event_clearance_group %in% c('TRAFFIC RELATED CALLS','FRAUD CALLS','BURGLARY', 'BIKE')) %>%
  ggplot(aes(district_sector, fill = event_clearance_group)) + geom_bar()
```

adding lables to graphs
```{r}
donor %>%
  ggplot(aes(primary_general, fill = primary_general)) +
  geom_bar () +
  labs(x = 'Election Type', y = 'Donations (#)')
```

now for the police data
```{r, warning=FALSE, message=FALSE}}
police %>% 
  filter(
    district_sector %in% c('B', 'E', 'D', 'R', 'O', 'C', 'K') &
          event_clearance_group %in% c('TRAFFIC RELATED CALLS','FRAUD CALLS','BURGLARY', 'BIKE')) %>%
  ggplot(aes(district_sector, fill = event_clearance_group)) + geom_bar() +
  labs(x = 'District Sector', y = 'Incidents by Event Clearance Group')
```

to remove the legend name:
```{r, warning=FALSE, message=FALSE}}
police %>% 
  filter(
    district_sector %in% c('B', 'E', 'D', 'R', 'O', 'C', 'K') &
          event_clearance_group %in% c('TRAFFIC RELATED CALLS','FRAUD CALLS','BURGLARY', 'BIKE')) %>%
  ggplot(aes(district_sector, fill = event_clearance_group)) + geom_bar() +
  labs(x = 'District Sector', y = 'Incidents by Event Clearance Group (#)', fill = element_blank())
```

```{r}
library(tidyverse)
```

adding titles and subtitles
```{r, warning=FALSE, message=FALSE}}
police %>% 
  filter(
    district_sector %in% c('B', 'E', 'D', 'R', 'O', 'C', 'K') &
          event_clearance_group %in% c('TRAFFIC RELATED CALLS','FRAUD CALLS','BURGLARY', 'BIKE')) %>%
  ggplot(aes(district_sector, fill = event_clearance_group)) + geom_bar() +
  labs(x = 'District Sector', y = 'Incidents by Event Clearance Group (#)', fill = element_blank(), title = 'Incidents by Clearance Events', subtitle = 'per District Sector')
```

to get themes
```{r, warning=FALSE, message=FALSE}}
police %>% 
  filter(
    district_sector %in% c('B', 'E', 'D', 'R', 'O', 'C', 'K') &
          event_clearance_group %in% c('TRAFFIC RELATED CALLS','FRAUD CALLS','BURGLARY', 'BIKE') ) %>%
  ggplot(aes(district_sector, fill = event_clearance_group)) + 
    geom_bar() +
    labs(x = 'District Sector', y = 'Incidents by Event Clearance Group (#)', fill = element_blank(), title = 'Incidents by Clearance Events', subtitle = 'per District Sector') + 
  theme_economist() + 
  scale_fill_economist()
```

build a histogram for the donor data
```{r}
donor %>%
  filter(amount <1000) %>%
  ggplot(aes(amount)) + geom_histogram()
```
now with density
```{r, warning=FALSE, message=FALSE}}
donor %>%
  filter(amount <1000) %>%
  ggplot(aes(amount)) + geom_density()
```

now try scatterplots
```{r, warning=FALSE, message=FALSE}}
donor %>%
  ggplot(aes(receipt_year, election_year)) + geom_point()
```

now line plot
```{r, warning=FALSE, message=FALSE}}
donor %>%
  ggplot(aes(receipt_year, amount)) + geom_line()
```

now make a scatterplot for police
```{r, warning=FALSE, message=FALSE}}
police %>%
  ggplot(aes(latitude, longitude, color = district_sector, alpha = district_sector)) + geom_point() +
  theme_classic()
```

aggregate data
```{r}
donor %>% group_by(type, party) %>% summarise(dollars = sum(amount, na.rm = TRUE))
```

turn it into a chart. **coord_flip turns bars sideways
```{r}
donor %>% 
  group_by(type, party) %>% 
  summarise(dollars = sum(amount, na.rm = TRUE)) %>%
  ggplot(aes(type, dollars, fill = party)) + geom_bar(stat = "identity") + 
  coord_flip() + 
  scale_y_continuous(labels = dollar) + theme_classic()
```

exercise
```{r}
donor %>%
  group_by(contributor_state) %>%
  summarise(avg_amount = mean(amount, na.rm = TRUE)) %>%
  filter(! contributor_state %in% NA) %>%
  ggplot(aes(reorder(contributor_state, -avg_amount), avg_amount)) +
  geom_bar(stat = 'identity') +
  scale_y_continuous(labels = dollar) + theme_classic()
```

```{r}
str(police)
```

lets use event clearance ampm and district sector
```{r, warning=FALSE, message=FALSE}}
police %>% 
  filter(
    district_sector %in% c('B', 'E', 'D', 'R', 'O', 'C', 'K') ) %>%
  ggplot(aes(district_sector, fill = event_clearance_ampm)) + 
    geom_bar() +
    labs(x = 'District Sector', y = 'Incidents by Time of Day', fill = element_blank(), title = 'Incidents by Time of Day', subtitle = 'per District Sector') + 
  theme_economist() + 
  scale_fill_economist()

```

```{r, warning=FALSE, message=FALSE}}
police %>% 
  filter(
          event_clearance_group %in% c('TRAFFIC RELATED CALLS','FRAUD CALLS','BURGLARY', 'BIKE') ) %>%
  ggplot(aes(event_clearance_group, fill = event_clearance_ampm)) + 
    geom_bar() + coord_flip() +
    labs(x = 'District Sector', y = 'Incidents by Time of Day', fill = element_blank(), title = 'Incidents by Time of Day', subtitle = 'per District Sector') + 
  theme_economist() + 
  scale_fill_economist()

```


